---
title: "P8105_hw2_asf2164"
output: github_document
date: "2022-09-30"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

##Problem 1 
```{r}
library(tidyverse)
library(readxl)

trans_ent =
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(
      Route8 = "c",
      Route9 = "c",
      Route10 = "c",
      Route11 = "c"
    )
  ) %>%
  janitor::clean_names() %>%
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    starts_with("route"),
    entry,
    exit_only,
    vending,
    entrance_type,
    ada
  ) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

trans_ent %>%
  select(station_name, line) %>%
  distinct

trans_ent %>%
  filter(ada == TRUE) %>%
  select(station_name, line) %>%
  distinct

trans_ent %>%
  filter(vending == "NO") %>%
  pull(entry) %>%
  mean

trans_ent %>%
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") %>%
  filter(route == "A") %>%
  select(station_name, line) %>%
  distinct

trans_ent %>%
  pivot_longer(route1:route11,
               names_to = "route_num",
               values_to = "route") %>%
  filter(route == "A", ada == TRUE) %>%
  select(station_name, line) %>%
  distinct
```


## Problem 2

```{r}
library(tidyverse)
library(readxl)

mrtrash_df =
  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls))) %>%
  mutate(across(.col = (c("year")), as.integer)) %>%
  mutate(source = "mr_trash", dumpster = as.numeric(dumpster))

proftrash_df =
  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(across(.col = (c("year")), as.integer)) %>%
  mutate(source = "prof_trash")

trash_tidy =
  bind_rows(mrtrash_df, proftrash_df) %>%
  janitor::clean_names() %>%
  select(-x15,-x16)

nrow(trash_tidy)
with(trash_tidy, sum(weight_tons[source == 'prof_trash'], na.rm = TRUE))
with(trash_tidy, sum(sports_balls[source == 'mr_trash' & year == "2020"], na.rm = TRUE))


```

There are `r nrow(trash_tidy)` observations. The key variables are weight_tons (weight of trash collected in tons) and sports_balls (number of sports balls collected). The total weight of trash collected by Professor Trash Wheel is `r with(trash_tidy, sum(weight_tons[source == 'prof_trash'], na.rm = TRUE))`. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r with(trash_tidy, sum(sports_balls[source == 'mr_trash' & year == "2020"], na.rm = TRUE))`. 

## Problem 3

```{r}
library(tidyverse)
monthly_pols_df =
  read_csv("data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day")) %>%
  mutate(across(.col = (c("year", "month")), as.integer)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(president = case_when(prez_gop == 1 ~ "gop", prez_dem == 1 ~ "dem")) %>%
  select(-prez_dem,-prez_gop,-day)

snp_df =
  read_csv("data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate(across(.col = (c("year", "month")), as.integer)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(year = if_else(year < 23, 2000 + year, 1900 + year)) %>%
  select(-day)

unemployment_df =
  read_csv("data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment percent") %>%
  mutate(across(.col = (c("year")), as.integer)) %>%
  mutate(
    month = case_when(
      month == "jan" ~ "Jan",
      month == "feb" ~ "Feb",
      month == "mar" ~ "Mar",
      month == "apr" ~ "Apr",
      month == "may" ~ "May",
      month == "jun" ~ "Jun",
      month == "jul" ~ "Jul",
      month == "aug" ~ "Aug",
      month == "sep" ~ "Sep",
      month == "oct" ~ "Oct",
      month == "nov" ~ "Nov",
      month == "dec" ~ "Dec"
    )
  )

snpmonths_df =
  left_join(monthly_pols_df, snp_df, by = c("year", "month"))

fivethirtyeight_df =
  left_join(snpmonths_df, unemployment_df, by = c("year", "month"))


nrow(fivethirtyeight_df)
ncol(fivethirtyeight_df)
max(fivethirtyeight_df$year, na.rm = TRUE)
min(fivethirtyeight_df$year, na.rm = TRUE)
```

The pols-month dataset contained 822 observation and 9 variables related to the number of national politicians who are democratic or republican at any given time. 

The snp dataset contained 787 observation and 2 variables related to Standard & Poor's stock market index, which is a representative measure of the stock market as a whole. 

The unemployment dataset contains 68 observation and 13 variables and shows the percentage of unemployment by month and year. 

The new dataset contains `r nrow(fivethirtyeight_df)` observations and `r ncol(fivethirtyeight_df)` variables. It has whether the president is democratic or republican (variable = president), the S&P stock market index (variable = close), and the unemployment percentage (variable = unemployment percentage) by year and month. The dataset spans from year `r max(fivethirtyeight_df$year, na.rm=TRUE)` to year`r min(fivethirtyeight_df$year, na.rm=TRUE)`. 

