---
title: "P8105_hw2_asf2164"
output: github_document
date: "2022-09-30"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 2

```{r}
library(tidyverse)
library(readxl)

mrtrash_df = 
 read_excel("data/Trash-Wheel-Collection-Totals-7-2020.xlsx", sheet = "Mr. Trash Wheel") %>% 
 janitor::clean_names() %>% 
 drop_na(dumpster) %>% 
mutate(sports_balls, as.integer(round(sports_balls))) %>% 
mutate(source = "mr_trash", dumpster = as.numeric(dumpster)) 

           
proftrash_df = 
 read_excel("data/Trash-Wheel-Collection-Totals-7-2020.xlsx", sheet = "Professor Trash Wheel") %>%
 janitor::clean_names() %>% 
 drop_na(dumpster) %>% 
mutate(sports_balls, as.integer(round(sports_balls))) %>% 
mutate(source = "prof_trash")

trash_tidy = 
  bind_rows(mrtrash_df, proftrash_df) %>% 
  janitor::clean_names() %>% 
  select(-x15, -x16, -x17)  
  
nrow(trash_tidy)
with(trash_tidy, sum(weight_tons[source == 'prof_trash']))
with(trash_tidy, mean(sports_balls[source == 'mr_trash'| year == "2020"]))


```

There are `nrow(trash_tidy)` observations. The key variables are weight_tons (weight of trash collected in tons) and sports_balls (number of sports balls collected). The total weight of trash collected by Professor Trash Wheel is `with(trash_tidy, sum(weight_tons[source == 'prof_trash']))`. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `with(trash_tidy, mean(sports_balls[source == 'mr_trash'| year == "2020"]))`. 

## Problem 3

```{r}
library(tidyverse)
monthly_pols_df = 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(col=mon, into=c("year", "month", "day")) %>%
  mutate(across(.col = (c("year", "month")), as.integer)) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(president = case_when(prez_gop==1 ~ "gop", prez_dem == 1 ~ "dem")) %>% 
  select(-prez_dem, -prez_gop, -day)  

snp_df = 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into=c("month", "day", "year")) %>% 
  mutate(across(.col = (c("year", "month")), as.integer)) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(year = if_else(year < 23, 2000+year, 1900+year)) %>% 
  select(-day)

unemployment_df = 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment percent"
  ) %>%
  mutate(across(.col = (c("year")), as.integer)) %>% 
  mutate(month = case_when(month == "jan" ~ "Jan", month=="feb" ~ "Feb", month=="mar" ~ "Mar", month=="apr" ~ "Apr", month=="may" ~ "May", month=="jun" ~ "Jun", month=="jul" ~ "Jul", month=="aug" ~ "Aug", month=="sep" ~ "Sep", month=="oct" ~ "Oct", month=="nov" ~ "Nov", month=="dec" ~ "Dec"))
  
snpmonths_df = 
  left_join(monthly_pols_df, snp_df, by = c("year", "month"))

fivethirtyeight_df =
  left_join(snpmonths_df, unemployment_df, by = c("year", "month"))


```



